---
title: "ch-2-some-basics"
author: "Sonya Hua"
date: "January 6, 2018"
output: html_document
---

```{r setup, include=FALSE}
library(knitr)
knitr::opts_chunk$set(echo = TRUE)

# set global display parameters
knitr::opts_chunk$set(fig.width=8, fig.height=5, echo = TRUE, fig.align="center") 
```

In order to combine multiple items of different types into a continuous output, use `cat`. `cat` puts a space between each item by default. You must provide a newline character ('\n') to terminate the line

```{r}
cat("The zero occurs at", 2*pi, "radians", "\n")

fib <- c(0,1,1,2,3,5,8,13,21,34)
cat("The first few Fibonacci numbers are:", fib)
```

### Listing Variables

Use `ls` to list variables and functions defined in workspace. It returns a vector of strings giving hte names of objects in workspace
```{r}
ls()
```
`ls.str` returns the variable name and applies the `str` function to the variable
```{r}

# For more detail
ls.str()
```

### Deleting variables & Erasing workspace

Use `rm` to remove variables or functions from workspace.
Use `rm(list=ls())` where list argument consts of a vector of names of variables to remove
```{r}
x <- "Hi"
rm(x)


# Erase all variables and functions in a workspace
rm(list=ls())
```

### Create a vector

A vector is a set of values of one data type in the form `c(.....)`

If the arguments of a vector contains vectors themselves, it flattens them and combines them into one single vector. 

```{r}
v1 <- c(1,2,3)
v2 <- c(4,5,6)

c(v1,v2)
```

### Basic Statistics

```{r}
x <- c(0,1,1,2,3,5,8,13,21,34)

mean(x)
median(x)
sd(x)
var(x)
```
To calculate correlation and covariance, there must not be any NA's. Even one missing value will cause the functions to return NA. To override this, use `na.rm==TRUE` to ignore NA's
```{r}
y <- log(x+1)

cor(x,y)
cov(x,y)
```
```{r}
x <- c(0,1,1,2,3,NA)
mean(x,na.rm=TRUE)
mean(x)
```
For data frames, `mean` and `sd` will calculate per column variable. 
```{r}
dframe <- data.frame(data(iris))
data(iris)

drop <- 'Species'
df <- iris[, !names(iris) %in% drop]

str(df)

# Returns a correlation matrix
cor(df)

# Returns a covarariance matrix
var(df)
```

The `median` function does not understand data frames. To calculate median of each column, use `apply` to each column with median funtion

### Creating Sequences

```{r}
0:5

seq(from=0, to=100, by=5)

# Repeated values
rep(1, times=5)

# Count backwards
10:0

# Specify a length for the output sequence and R will automatically calculate the increments needed
seq(from=0,to=20, length.out=7)

```

### Vector Arithmetic

Arithmetic on vectors is on an element wise level. Length of each vector must be the same
```{r}
v <- c(11,12,13,14,15)
w <- c(1,2,3,4,5)

v+w
v-w
v*w
v/w
w*v

# Calculdate z-score manually
(w - mean(w))/sd(w)
```

For lists, one generally uses [[ to select any single element, whereas [ returns a list of the selected elements.

The [[ form allows only a single element to be selected using integer or character indices, whereas [ allows indexing by vectors.

### Defining functions

Use `function` keyword followed by a list of parameters and the function body. `{ }` should be used around the body function for expressions.
```{r}
# To calculate coefficient of variation

cv <- function(x) sd(x)/mean(x)  # create a function and assign it to cv
cv(1:10) # use 1:10 as value of x
```

A multiline function requires curly braces to delimit the start and end of the function.

```{r}
# Cacluate Euclid's algorithm for computing the greatest common divisor of 2 integers
gcd <- function(a,b) {
  if (b==0) return (a)
  else return(gcd(b, a %% b))
}

```
`%%` modulo operator 

R also allows anonymous or undefined functions. 

```{r}
lst <- c(1,2,3,3,4,5)
lapply(lst, function(x) sd(x) / mean(x))
```
####Other tips

`>` = normal prompt
`+` = continuation prompt

Use `==` for comparisons and `=` for assigning variables

Use `library` or `require` to load packages into search path

If a variable `lst` contains a list, it can be indexed in 2 ways: `lst[[n]]` is the nth element of hte list; whereas `lst[n]` is a list whose only element is the nth element of `lst`. 

Do not pass multiple arguments to a single-argument function. Functions like `mean` only takes one argument and if we want to calculate the mean of 9, 10, and 11, we need to do `mean(c(9,10,11))`