---
title: "ch-3-navigating-r"
author: "Sonya Hua"
date: "January 6, 2018"
output: html_document
---

```{r setup, include=FALSE}
library(knitr)
knitr::opts_chunk$set(echo = TRUE)

# set global display parameters
knitr::opts_chunk$set(fig.width=8, fig.height=5, echo = TRUE, fig.align="center") 
```

### Getting and setting the working directory

```{r}
getwd()

#setwd() to set working directory

```

### Saving Workspace

The workspace holds R variables and functions. It's created when R starts. Use `save.image` function to save worksapce momentarily. The worksapce is written to a filecalled .Rdata in the working directory. When R starts, it looks for htat file and initialized the wrokspace. 

```{r}
# To see list of packages currently loaded into R
search()

# To access built-in datasets
data(iris)
head(iris)
help(iris)  # Brings up a description of the data set

# To see a table of contents for datasets
data()
data(package="MASS")

# To view list of installed packages
library()
```


### Setting a default CRAN mirror

You want to set a default CRAN mirror so R does not prompt for one everytime

```{r}
chooseCRANmirror()
```

### Configuring default packages to be loaded each time R runs

Set a config parameter called `defaultPackages`, a list of packages loaded by R at startup. It initally contains a system-deflined list. If I append package names to the list then R will load them, saving me from having to call `library` each time. Include this snippet in your r profile.

R executes profile scripts when it starts, freeing you from repeatedly loading often-used packages or tweaking the R configuration options

```{r}

### Loading packages at start
pkgs <- getOption("defaultPackages")
pkgs <- c(pkgs, "car", "dplyr", "tidy", "ggplot2", "RColorBrewer", "plyr")
options(defaultPackages=pkgs)
rm(pkgs)

```


