---
title: "ch-5-data-structures"
author: "Sonya Hua"
date: "January 7, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Covering matrices, lists, factors, and data frames.

Vectors are homogeneous, can be indexed by position, can have names.
Lists are heterogeneous, can be indexed by position, lets you extract sublists, and can have names. 

Mode: how it is stored in memory (i.e. char, num, etc.)
Class: defines the object's abstract type (i.e. a Date object consists of a single nnumber but has a class of Date)
Scalars: a vector that contains exactly 1 element
Matrices: a vector that has dimensions (i.e. 2x3)
```{r}
A <- 1:6
dim(A) <- c(2,3)
A
```

Arrarys: 3D or n-dimensional structures
```{r}
D <- 1:12
dim(D) <- c(2,3,2)
D
```

Factors: has levels whic is used for efficient storage of data and can be used in statistical models. There are 2 types: Categorical variables and grouping

Data Frames: tabular data structure with rows and columns. Data frame is a list with elements and vectors. 

```{r}
# To insert data into a vector

append(1:10, 99999, after = 5)
```

```{r}
# To create a factor
v <- c(1,2,3,4,5)
f <- factor(v)
f

# Create levels
f <- factor(v, c("Strongly DA","Somewhat DA","Neutral","Somewhat A", "Strongly A"))
f

# To convert variable into a factor, just call factor on the variable
```

When an R function requires a factor, it usually converts data to a factor automatically. 

```{r}
# To combine vectors into one large vector and create a parallel factor var that identifies each value's original group

comb <- stack(list(v1=values1, v2=values2, v3=values3))
# V1, v2, and v3 are the values in 2nd column
```

For lists ,we can access single elements using double brackets

```{r}
years <- list(c(1999,2000), 2007, 1989, 1991, 2011)
years


```


```{r}

# This is an element, not a list. It is the nth element of years
years[[1]]
class(years[[1]])
```
```{r}

# This is a list, not an element. The list contains 1 element taken from the nth element of list
years[1]
class(years[1])
```
The difference becomes apparent when we inspect the structure of hte result - one is numeric, the other is a list

```{r}
cat(years[[1]])
```

```{r}

# Error since it's not atomic
cat(years[1])
```

### Slecting List Elements by Name
```{r}
# Access list elements by names

years <- list(Kennedy=c(1960, 1961), Johnson=1964, Carter=1976, Clinton=1994)
years[["Kennedy"]] # Element
years["Kennedy"] # List
years$Kennedy
```

### Building a Name and Value list

You can associate the names and values by creating an emptry list then populating with a vectorized assignment statement

```{r}
# Sometimes you have a vector of names an dvector of values to go along with it
(values <- pnorm(-2:2))
names <- c("A","B","C","D","E")

alist <- list()
alist[names] <- values
alist
```

```{r}
for (nm in names(alist))
  cat("The", nm, "is", alist[[nm]], "\n")
```

```{r}
# To remove a list element assign NULL
years[["Johnson"]] <- NULL

# To unlist into a vector for basic stat functions that only work on vctors
unlist(years)   
```



### To remove null elements from a list

`sapply` is equivalent to `lapply` but works for vectors and amtrices

```{r}
alist2 <- list("Moe",NULL, "Sal")
alist2

# Removes NULL element from list
alist2[sapply(alist2, is.null)] <- NULL
alist2


# Or

alist2[is.na(alist2)] <- NULL
```

### Create a data frame
```{r}
dfrm <- data.frame(v1, v2, v3, f1, f2)

```
### Selecting Rows and Columns of DF more easily

```{r}
data(AMSsurvey)
ams <- data("AMSsurvey")
AMSsurvey
subset(AMSsurvey, sex=="Female")
subset(AMSsurvey, select=citizen)
```

```{r}

# combine select and subset arguments

subset(AMSsurvey, select=c(sex,citizen), subset=(sex=="Female" & citizen=="US"))
```
```{r}
sample <- AMSsurvey
colnames(AMSsurvey)

```


```{r}
# To remove rows with NAs, use na.omit
sample <- na.omit(AMSsurvey)


# To exclude columns by name

subset(sample, select = -sex)

# To exclude  column by position
sample[-1]
```

```{r}
# To combine data frames by rows
allrows <- rbind(df1, df2)
allcolumns <- cbind(df1, df2)
```

```{r}
# To access data frame columns more easily instead of repeatedly typing the df name

attach(sample)
(sex)
detach(sample)
```

## Converting data structures into another type

```{r}
as.data.frame(x)
as.llist(x)
as.matrix(x)
as.vector(x)
```

