---
title: "ch-6-strings-and-dates"
author: "Sonya Hua"
date: "January 7, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Date: a calendar date but not a clock time.
POSIXct: a datetime class that can represent a moment in time with an accuracy of 1 second. It's stored as the number of seconds since 1/1/1970. Recommended format for storing datetime info in data frame. 
POSIXlt: a datetime class stored as year, month, day, hour, minute, second

timeDate package is a high-powered package with well-thought-out facilities for handling dates and times, including date arithmetic, business days, holidays, conversions, and generalized handling of time zones. 

### Getting the length of a string

```{r}
# count the number of characters in a strong or vector
s <- c("Moe","Larry", "Curly")
nchar(s)

# Count the number of elements
length("Moe")
length(s)
```

### Concatenating STrings

```{r}

# By default, paste inserts a single space between pairs of strings. 
paste("Everybody", "loves","stats")
paste("Everybody", "loves", "states", sep=".")

# It converts an numbers to characters
paste("The square root of pi is", sqrt(pi))
```

```{r}
# To extract a portion of a string according to position, use substr(string,start,end)

substr("Statistics", 1, 4)
substr("Statistics", 7, 10)

# It can be aplied to every element of a string

ss <- c("Jeff", "Sonya", "Amy")
substr(ss,1,1)

cities <- c("New York, NY", "Los Angeles, CA", "Peoria, IL")
substr(cities, nchar(cities)-1, nchar(cities))
substr(cities, 1, nchar(cities)-4)
```
### Split a string according to a delimiter
```{r}
path <- "/home/mike/data/trials.csv"
strsplit(path, "/")
```

### Replacing substrings

```{r}

# To replace first instance of a substring, use sub(find, replace, string)
s <- "Curly is the smart one. Curly is funny, too."
s <- sub("Curly", "Moe", s)
sub("Moe", "Sonya", s)

# To replace all instances of a substring, use gsub(old, new, string)
s <- "Curly is the smart one. Curly is funny, too."
gsub("Curly", "Sonya" ,s)

# To remove a substring for once, set the substring to be empty
sub("the ","", s)

```

# Generate pairwise combinations of strings

The outer function can accept any function for its third argument. In this recipe, we replace multiplication with string concatenation (paste). The results is a combination of all strings


`outer(string1, string2, function, sep="")`
```{r}
# Use outer and paste functions together to generate the matrix of all possible combinations
locations <- c("NY","LA", "CHI", "HOU")
treatments <- c("T1","T2","T3")

outer(locations, treatments, paste, sep=".")
dput(outer(locations, treatments, paste, sep="."))

# To output as a vector instead
as.vector(outer(locations, treatments, paste, sep="."))
```

Suppose we want to only output unique pairwise combinations of treatments. We can eliminate redundancy by removing the lower triangle (or upper triangle). The `lower.tri` function identifies that triangle, so inverting it identifies all elements outside the lower triangle

```{r}
m <- outer(treatments, treatments,paste,sep='.')
m
lower.tri(m) # captures values of lower triangle or the repetitive values
m[!lower.tri(m)] # removes lower triangle values
```

### Get the current date
```{r}
Sys.Date()
class(Sys.Date())
```

### Convert a string into a date

How to convert "2010-12-31" into a date? We can use `as.Date` but we must know the format of the string. By default, `as.Date` assuems the string looks like yyyy-mm-dd. The string can be in other date formats but we then must specify the `format=` argument so it can be interpreted

```{r}
as.Date("2017-12-31")
as.Date("12/31/2017", format="%m/%d/%Y")
```

### Convert Year, Month, and Day into a Date

```{r}
#ISOdate(year, month, day)

ISOdate(2012,2,29)
```
### Extract parts of a date

Convert a `Date` object to a `POSIXlt`, a list of date parts. Extract date parts such as the day of week, day of year, calendar day, calendar month, or calendar year

```{r}
d <- as.Date("2010-03-15")
p <- as.POSIXlt(d)

p$mday  # Day of Month
p$mon  # Month
p$year  # year
p$wday  # weekday (0 = Sunday)
```

### Create a Sequence of Dates

```{r}
s <- as.Date("2012-01-01")
e <- as.Date("2012-02-01")

seq(from=s, to=e, by=1)  # by 1 day

seq(from=s, by=1, length.out=7)
```

```{r}
seq(from=s, by="month", length.out=12)
```
```{r}
seq(from=s, by="week", length.out=52)
```

