---
title: "ch-8-probability"
author: "Sonya Hua"
date: "January 10, 2018"
output: html_document
---

Distrete distributions include:
- binomial
- geometric
- hypergeometric
- negative binomial
- poisson

Continuous distributions include:
- beta
- cauchy
- chi-squeared
- exponential
- F
- gamma
- log-normal
- logistic
- student's T
- normal
- uniform
- Weibull
- Wilcoxon

### Counting the number of combinations

To calculate the number of combinations of *n* items taken *k* at a time, use `choose`. This calculates the number of distinct subsets of size k that can be created from *n* items

```{r}

# How many ways can we select 3 items from 5 items?
choose(5,3)

# How many ways can we select 3 items from 50 items?
choose(50,3)
```

### Generate Combinations of items

```{r}

# Generate all combinations of 1,2,3 taken twice at a time
combn(c(1,2,3), 2)

# Generate combinations of strings
combn(c("T1","T2","T3","T4","T5"),3)
```

### Generate random numbers 

```{r}
# Generate a uniform random number between 0 and 1
runif(1)
runif(5)

# Generate random numbers from other distributions. For a given distribution, the name of the random number generator is "r" prefixed to the distribution.

rnorm(1)
rchisq(1,df=3)

```

```{r}
runif(10)
```

```{r}
rbinom(10, size=100, prob=0.5)
```

```{r}
# Calculate 100 draws of a normal variate whose mean is itself randomly distributed and with hyperparameters of mean = 0 , sd = 0.2

means <- rnorm(100, mean=0, sd=0.2)
rnorm(100, mean=means, sd=1)
```

### Generate Reproducible Random Numbers

```{r}
set.seed(123)
runif(10)


runif(10)

set.seed(123)
runif(10)
```

### Generating a Random Sample